Generalized Tractability

### 2023-04-28 GeneralizedTractability_230418.tex
Updates from Overleaf

This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent.  The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering.

@@

@@

\begin{theorem}\label{thm_main_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable iff there exists $\vp\ge \vzero$ and a positive integer $L_{\vp}$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let
\[
\widetilde{\cp}_{\textup{opt}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \text{ and }  \eqref{eq:tractiff4} \text{ does not hold } \forall \vp \in [\vzero,\vp^*/2)\}.
\]
Then $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.
\
\end{theorem}

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       + \frac{1}{T(\varepsilon^{-1},d, \vp)} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}  \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right] \\
       & =\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + \left [  S_{\vp} + \frac{1}{T(0,1, \vp)}\right] T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) +\left [  S_{\vp} + \frac{1}{T(0,1, \vp)}\right] T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)}  T(\varepsilon^{-1},d,2\vp)  +\left [  S_{\vp} + \frac{1}{T(0,1, \vp)}\right] T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp) \qquad \text{by \eqref{eq:Tnondecreasing}}.
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

>>>>

This means that the non-negative $\lambda_{i,d}$ are the singular values of the operator $\SOL_d$. We remark that in the literature on Information-Based Complexity the singular values of $\SOL_d$ are often denoted by $\sqrt{\lambda_{i,d}}$ instead of $\lambda_{i,d}$. However, since we assume non-negativity of the $\lambda_{i,d}$, the notation used here is equivalent. 
\peter{[PETER: I am wondering whether we should go back to the more common notation once the manuscript is finalized (not right away). It would make the paper easier to read for the IBC community.]}

The non-decreasing ordering of the eigenvalues for each $d$ facilitates the proof of conditions for tractability, but our conditions do not depend on the ordering.

@@

Now we take this upper bound further, specializing to the case of $d'=1$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \sum_{i=1}^{L_{\vp}-1} \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}
       + \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
       + \frac{1}{T(\varepsilon^{-1},1, \vp)} \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{(L_{\vp}-1)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right]}_{=:C_{\vp}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}

\peter{
[PETER: This is an alternative part of the proof. If it is correct we could use it instead of the black text.

Let $\widetilde{J}\in\naturals$. Since the $\lambda_{i,d}$ are non-increasing in $i$, it follows that the $T(\lambda_{i,d}^{-1},d',\vp)$ are non-decreasing in $i$ for any $d'\in\naturals$. In particular, for all $n\ge \widetilde{J}$, we have
\begin{align*}
    \lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{\widetilde{J},d}
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},d',\vp)} \le \cdots \le \frac{1}{T(\lambda_{\widetilde{J}, d}^{-1},d',\vp)} \\
     \implies \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp) }
    \le &\frac{1}{n-\widetilde{J}+1} \sum_{i=\widetilde{J}}^n  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) }
    \le \frac{1}{n-\widetilde{J}+1} \sum_{i=\widetilde{J}}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)}.
\end{align*}

Thus, we can conclude from the previous line that
\begin{multline} \label{eq:implicate_a}
    n - \widetilde{J} +1 \ge T(\varepsilon^{-1},d',\vp) \sum_{i=\widetilde{J}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp)} \\
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},d',\vp)} \le
   \frac{1}{n-\widetilde{J}+1} \sum_{i=\widetilde{J}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d',\vp) } \le \frac{1}{T(\varepsilon^{-1},d',\vp)} \qquad \forall \varepsilon > 0, \ d,d' \in \naturals.
\end{multline}

Now we take this upper bound further, specializing to the case of $d'=1$ and $\widetilde{J}=L_{\vp}$:
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},1, \vp) \Biggl [ \underbrace{\sum_{i=L_{\vp}}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S_{\vp} <\infty\ \text{by \eqref{eq:strong_tractiff3}} }
       + \frac{L_{\vp}}{T(\varepsilon^{-1},1, \vp)} \Biggr] \\
       & \le T(\varepsilon^{-1},1,\vp) \underbrace{\left[S_{\vp} + \frac{L_{\vp}}{T(0,1, \vp)}\right]}_{=:C_{\vp}}\\
       & =  C_{\vp}T(\varepsilon^{-1},1,\vp).
%       & \le \underbrace{2 C_{1,\vp}}_{\text{since } T(\varepsilon^{-1},1,\vp) \ge 1} T(\varepsilon^{-1},1,\vp)
%       = \underbrace{C_{\vp}}_{:=2 C_{1,\vp}}T(\varepsilon^{-1},1,\vp).
\end{align*}

END OF THE ALTERNATIVE PROOF.]

}

This means that we have strong $T$-tractability via \eqref{eq:strong_tract_def}, and verifies the sufficiency of \eqref{eq:strong_tractiff3}.

@@

\begin{theorem}\label{thm_main_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable iff there exists $\vp\ge \vzero$ and a positive integer $L_{\vp}$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let
\[
\widetilde{\cp}_{\textup{opt}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \text{ and }  \eqref{eq:tractiff4} \text{ does not hold } \forall \vp \in [\vzero,\vp^*/2)\}.
\]
Then $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.

\end{theorem}

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       + \frac{1}{T(\varepsilon^{-1},d, \vp)} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}  \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right] \\
       & =\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + \left [  S_{\vp} + \frac{1}{T(0,1, \vp)}\right] T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) +\left [  S_{\vp} + \frac{1}{T(0,1, \vp)}\right] T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)}  T(\varepsilon^{-1},d,2\vp)  +\left [  S_{\vp} + \frac{1}{T(0,1, \vp)}\right] T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp} + \frac{1}{T(0,1, \vp)}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp) \qquad \text{by \eqref{eq:Tnondecreasing}}.
\end{align*}

\peter{
[PETER: This is an alternative part of the proof. If it is correct we could use it instead of the black text.

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$. By the argument in the proof of the sufficient condition for Theorem \ref{thm_main_strong_tract2}, it follows from \eqref{eq:compUP} with $d'=d$ that the information complexity has an upper bound of
\begin{equation*}
	 \comp(\varepsilon,d) \le  T(\varepsilon^{-1},d,\vp) \sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)} + J(\vp,d)
\end{equation*}
 Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & \le T(\varepsilon^{-1},d,\vp) \Biggl [ 
       \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}}
       + \frac{J(\vp,d)}{T(\varepsilon^{-1},d, \vp)} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ S_{\vp} + \frac{L_{\vp} T(0,d,\vp)+1}{T(\varepsilon^{-1},d, \vp)}\right]\\
       & \le T(\varepsilon^{-1},d,\vp)\left[S_{\vp} +\frac{L_{\vp} T(0,d,\vp)}{T(\varepsilon^{-1},d, \vp)}
       + \frac{1}{T(0,1, \vp)}\right]
       \qquad \text{by \eqref{eq:Tnondecreasing}}   \\
       & \le T(\varepsilon^{-1},d,\vp) 
       \underbrace{\left [  S_{\vp} + L_{\vp} + \frac{1}{T(0,1, \vp)}\right]}_{=:C_{\vp}}
       \qquad \text{by \eqref{eq:Tnondecreasing}}  
\end{align*}

END OF THE ALTERNATIVE PROOF.]
}

It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.
